\bf Nonparametric Block Bootstrap Kolmogorov-Smirnov Goodness-of-Fit Test

### 2025-05-25 manuscript/manuscript.tex

The standard one-sample Kolmogorov--Smirnov (KS) test is widely
recognized as an effective good-of-fit test for continuous distributions.
Consider $X_1, ..., X_n$, a random sample of size~$n$ from some continuous
distribution and the null hypothesis $H_0$ that $X_i$'s follow a specific
hypothesized distribution~$F(x; \theta_0)$, where $\theta_0$ are the
specified parameters
of the hypothesized distribution family.
If we let $F_n(t) = \sum_{i=1}^n \mathbf{1}(\{X_i \le t\} / n$ be the empirical cumulative
distribution function of the sample, where $\mathbf{1}\{\cdot\}$ is the indicator
function, the KS test statistic takes the form
\[
  D_n = \sqrt{n} \sup_x | F_{n}(x) - F(x; \theta_0)|.
\]
As sample size $n\to \infty$, the distribution of $D_n$ converges to that of the
absolute value of standard Brownian bridge, which is known as the Kolmogorov
distribution \citep{stephens1974edf}. This distribution function is
precisely computable using contemporary statistical software
\citep{marsaglia2003evaluating}. The versatility of the KS test allows its
application across various domains, such as analyzing cosmic microwave
background radiation \citep{naess2012application}, monitoring the count rate of
radioactive data \citep{aslam2020introducing}, gear condition monitoring
\citep{andrade2001gear}, studying images of breast cancer tumors
\citep{demidenko2004kolmogorov}, and examining financial markets
\citep{lux2001turbulence}.

@@

First, let us review how the KS statistic is computed for an independent
sample with fitted parameters. Let $\hat\theta_n$ denote the parametrically
fitted parameters, which can be obtained from any consistent estimator with
asymptotic normality property.  Let $F_n$ be the empirical distribution
function based on $X_1,...,X_n$.
Define
\begin{equation*}
Y_n(x; \hat\theta_n) = \sqrt{n}(F_n(x) - F(x; \hat\theta_n)).
\end{equation*}
Then, the one-sample KS goodness of fit statistic is
\begin{equation*}
T_n := \sup_x|Y_n(x; \hat\theta_n)|.
\end{equation*}

@@

In summary, the procedure of the NPB KS test for an independent sample is
summarized as follows. Repeat the following steps for $b \in \{1, ..., B\}$.
\begin{enumerate}
\item
  Generate $X^{(b)}_1,...,X^{(b)}_n$ by sampling $X_1, \ldots, X_n$
  with replacement.
\item
  Obtain parametrically
  fitted parameters
	$\hat\theta^{(b)}_n$ of $\theta$ from $X^{(b)}_1,...,X^{(b)}_n$.
\item
  Obtain the empirical distribution function $F^{(b)}_n$ of
  $X^{(b)}_1,...,X^{(b)}_n$.
\item
  Calculate bootstrap KS statistic
  \[
    T^{(b)}_n = \sup_x \Big\vert \sqrt{n}\left(F^{(b)}_n(x)
    - F(x; \hat\theta^{(b)}_n)\right) - C_n(x) \Big\vert.
  \]
  where
  \[
    C_{n}(x) = \sqrt{n}(F_n(x) - F(x; \hat\theta_n))
  \]
  is the estimated bias term.
\end{enumerate}

@@

We now consider the case where $X_i$'s are realizations from a time series and
$X^{(b)}_1,...,X^{(b)}_n$ are generated by block bootstrap for
$b \in \{1, \ldots, B\}$.
Circular block-bootstrap can be done with overlapping blocks that can
wrap around the end of the series \citet{romano1992circular}.
Define blocks (assuming block size $l > 1$) as:
\begin{equation*}
Z_j =
    \begin{cases}
        \{X_j, \ldots, X_{j + l - 1}\}, & j = 1, \dots, n - l + 1,\\
        \{X_j, \ldots, X_n, X_1, \ldots, X_{j-n+l-1}\}, & j = n - l
        + 2 ,\dots, n.
    \end{cases}
\end{equation*}
A common
function for block size that is considered optimal is
$l = \lceil n^{1/3} \rceil$ \citep{buhlmann1999block}... \citet{hall1995blocking}
and
\citet{politis2004automatic} note that
the selection of block size should ideally take into account the strength
and direction of serial dependence. The plug-in approach for circular
bootstrap introduced in \citet{politis2004automatic}...
Now we draw $k$ blocks from the $(n - l + 1)$ blocks
of $Z_j$'s with replacement and then align them in the order they were picked to
form a block bootstrap sample. If $n$ is not a multiple of~$l$, the last block
selected will be reduced in size so that the final size of the block bootstrap
sample is $n$.

@@

In summary, the procedure of the NPBB test is
summarized as follows. Repeat the following steps for $b \in \{1, ..., B\}$.
\begin{enumerate}
\item
  Generate $X^{(b)}_1,...,X^{(b)}_n$ by applying circular block bootstrap
  on the original sample as
  defined previously.
\item
  Obtain parametrically fitted parameters
  $\hat\theta^{(b)}_n$ of $\theta$ from $X^{(b)}_1,...,X^{(b)}_n$.
\item
  Obtain the empirical distribution function $F^{(b)}_n$ of
  $X^{(b)}_1,...,X^{(b)}_n$.
\item
  Calculate bootstrap KS statistic
  \[
    T^{(b)}_n = \sup_x \Big\vert \sqrt{n}\left(F^{(b)}_n(x)
    - F(x; \hat\theta^{(b)}_n)\right) - K_n(x) \Big\vert.
  \]
  where
  \[
    K_{n}(x) = \sqrt{n}(\mathbb{E}^{*}[F^{(b)}_n(x)] -
    F(x; \theta_0^*))
  \]
  is the bias term, and
  the expected values $\mathbb{E}^{*}[F^{(b)}_n(x)]$ and
$\theta_0^* = \mathbb{E}^{*}[\hat\theta^{(b)}_n]$ can be
approximated by, respectively,
$\mathbb{E}_B^{*}[F^{(b)}_n(x)] = \frac{1}{B}\sum_{b = 1}^B F^{(b)}_n(x)$ and
$\mathbb{E}_B^{*}[\hat\theta^{(b)}_n]  =  \frac{1}{B}\sum_{b = 1}^B\hat\theta^{(b)}_n$.
\end{enumerate}
The p-value of the block bootstrap KS test can be approximated
as $p = \sum_{b=1}^B \mathbf{1}\{T^{(b)}_n > T_n\} / B$.

@@

\section{Conclusions}\label{sec:conclusion}

@@

>>>>

The standard one-sample Kolmogorov--Smirnov (KS) test is widely
recognized as an effective good-of-fit test for continuous distributions.
Consider $X_1,  \cdots , X_n$, a random sample of size~$n$ from some continuous
distribution and the null hypothesis $H_0$ that $X_i$'s follow a specific
hypothesized distribution~$F(x; \theta_0)$, where $\theta_0$ are the
specified parameters
of the hypothesized distribution family.
If we let $F_n(t) = \sum_{i=1}^n \mathbf{1}(\{X_i \le t\} / n$ be the empirical cumulative
distribution function of the sample, where $\mathbf{1}\{\cdot\}$ is the indicator
function, the KS test statistic takes the form
\[
  D_n = \sqrt{n} \sup_x | F_{n}(x) - F(x; \theta_0)|.
\]
As sample size $n\to \infty$, the distribution of $D_n$ converges to that of the
absolute value of standard Brownian bridge, which is known as the Kolmogorov
distribution \citep{stephens1974edf}. This distribution function is
precisely computable using contemporary statistical software
\citep{marsaglia2003evaluating}. The versatility of the KS test allows its
application across various domains, such as analyzing cosmic microwave
background radiation \citep{naess2012application}, monitoring the count rate of
radioactive data \citep{aslam2020introducing}, gear condition monitoring
\citep{andrade2001gear}, studying images of breast cancer tumors
\citep{demidenko2004kolmogorov}, and examining financial markets
\citep{lux2001turbulence}.

@@

First, let us review how the KS statistic is computed for an independent
sample with fitted parameters. Let $\hat\theta_n$ denote the parametrically
fitted parameters, which can be obtained from any consistent estimator with
asymptotic normality property.  Let $F_n$ be the empirical distribution
function based on $X_1, \cdots ,X_n$.
Define
\begin{equation*}
Y_n(x; \hat\theta_n) = \sqrt{n}(F_n(x) - F(x; \hat\theta_n)).
\end{equation*}
Then, the one-sample KS goodness of fit statistic is
\begin{equation*}
T_n := \sup_x|Y_n(x; \hat\theta_n)|.
\end{equation*}

@@

In summary, the procedure of the NPB KS test for an independent sample is
summarized as follows. Repeat the following steps for $b \in \{1,  \cdots , B\}$.
\begin{enumerate}
\item
  Generate $X^{(b)}_1, \cdots ,X^{(b)}_n$ by sampling $X_1, \ldots, X_n$
  with replacement.
\item
  Obtain parametrically
  fitted parameters
	$\hat\theta^{(b)}_n$ of $\theta$ from $X^{(b)}_1, \cdots ,X^{(b)}_n$.
\item
  Obtain the empirical distribution function $F^{(b)}_n$ of
  $X^{(b)}_1, \cdots ,X^{(b)}_n$.
\item
  Calculate bootstrap KS statistic
  \[
    T^{(b)}_n = \sup_x \Big\vert \sqrt{n}\left(F^{(b)}_n(x)
    - F(x; \hat\theta^{(b)}_n)\right) - C_n(x) \Big\vert.
  \]
  where
  \[
    C_{n}(x) = \sqrt{n}(F_n(x) - F(x; \hat\theta_n))
  \]
  is the estimated bias term.
\end{enumerate}

@@

We now consider the case where $X_i$'s are realizations from a time series and
$X^{(b)}_1, \cdots ,X^{(b)}_n$ are generated by block bootstrap for
$b \in \{1, \ldots, B\}$.
Circular block-bootstrap can be done with overlapping blocks that can
wrap around the end of the series \citet{romano1992circular}.
Define blocks (assuming block size $l > 1$) as:
\begin{equation*}
Z_j =
    \begin{cases}
        \{X_j, \ldots, X_{j + l - 1}\}, & j = 1, \dots, n - l + 1,\\
        \{X_j, \ldots, X_n, X_1, \ldots, X_{j-n+l-1}\}, & j = n - l
        + 2 ,\dots, n.
    \end{cases}
\end{equation*}
A common
function for block size that is considered optimal is
$l = \lceil n^{1/3} \rceil$ \citep{buhlmann1999block}.
\citet{hall1995blocking} and
\citet{politis2004automatic} note that
the selection of block size should ideally take into account the strength
and direction of serial dependence. The plug-in approach for circular
bootstrap introduced in \citet{politis2004automatic}.
Now we draw $k$ blocks from the $(n - l + 1)$ blocks
of $Z_j$'s with replacement and then align them in the order they were picked to
form a block bootstrap sample. If $n$ is not a multiple of~$l$, the last block
selected will be reduced in size so that the final size of the block bootstrap
sample is $n$.

@@

In summary, the procedure of the NPBB test is
summarized as follows. Repeat the following steps for $b \in \{1,  \cdots , B\}$.
\begin{enumerate}
\item
  Generate $X^{(b)}_1, \cdots ,X^{(b)}_n$ by applying circular block bootstrap
  on the original sample as
  defined previously.
\item
  Obtain parametrically fitted parameters
  $\hat\theta^{(b)}_n$ of $\theta$ from $X^{(b)}_1, \cdots ,X^{(b)}_n$.
\item
  Obtain the empirical distribution function $F^{(b)}_n$ of
  $X^{(b)}_1, \cdots ,X^{(b)}_n$.
\item
  Calculate bootstrap KS statistic
  \[
    T^{(b)}_n = \sup_x \Big\vert \sqrt{n}\left(F^{(b)}_n(x)
    - F(x; \hat\theta^{(b)}_n)\right) - K_n(x) \Big\vert.
  \]
  where
  \[
    K_{n}(x) = \sqrt{n}(\mathbb{E}^{*}[F^{(b)}_n(x)] -
    F(x; \theta_0^*))
  \]
  is the bias term, and
  the expected values $\mathbb{E}^{*}[F^{(b)}_n(x)]$ and
$\theta_0^* = \mathbb{E}^{*}[\hat\theta^{(b)}_n]$ can be
approximated by, respectively,
$\mathbb{E}_B^{*}[F^{(b)}_n(x)] = \frac{1}{B}\sum_{b = 1}^B F^{(b)}_n(x)$ and
$\mathbb{E}_B^{*}[\hat\theta^{(b)}_n]  =  \frac{1}{B}\sum_{b = 1}^B\hat\theta^{(b)}_n$.
\end{enumerate}
The p-value of the block bootstrap KS test can be approximated
as $p = \sum_{b=1}^B \mathbf{1}\{T^{(b)}_n > T_n\} / B$.

@@

\section{Discussion}\label{sec:conclusion}

@@

The choice of block size is well known to influence the performance of
block bootstrap methods, particularly for inference on dependent
data. As pointed out by \citet{hall1995blocking} and
\citet{lahiri1999theoretical}, the optimal block length depends on the
strength of serial dependence, and using a uniform choice across
different dependence settings can lead to size distortion. Several
data-driven procedures for selecting block size have been proposed,
including plug-in and cross-validation methods
\citep{buhlmann2002bootstraps, politis2004automatic}. A more
comprehensive treatment is given in
\citet{lahiri2013resampling}. Following these suggestions, we applied
the automatic block length selection method of
\citet{politis2004automatic} for circular block bootstrap in our
context. As reported in the Appendix, the method frequently selected
block sizes that were overly large—sometimes exceeding the sample
size—especially under strong serial dependence. This reduced the
variability of the bootstrap replicates and affected the reliability
of the test. These observations highlight that while automatic methods
are attractive in principle, their performance in small to moderate
samples and in goodness-of-fit settings like ours may be limited.
